---
title: "Aides Drug"
author: "David Barton"
date: "September 2025"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

For this assignment, I noticed the problem was in Categorical categorical. So the helper file Categorical Categorical was the best for this. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
females <- filter(triUmp,gender=="F")
```

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

After examining the chart, it was found that more people took the drug than those who took the placebo. Most of the people who got the virus took the drug. Showing that the drug isn't working as it should. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```

Out of the people who have aids, 3/5 who have are taking the drug have aids. Showing even more evidence, this time numerical, that the drug is not working correctly against aids. 58.82% of people took the drug and still have aids. 

## Inferential Results



<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

After examining Fisher's Exact Test for Count Data, the P-value is 6.596e-05, and the odds ratio is 1.543902. It is clear that it is the best option not to take the drug, as it has been shown that there is a higher chance of getting Aids while taking the medication.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


